Building and Checking Suffix Array Using Induced Sorting Method

### 2017-08-05 new-draft/draft.tex
2017/08/05

Suffix array~(SA) can be built by the induced sorting~(IS) method on both internal and external memory models. We propose two methods that enable any IS suffix sorter to build and check an SA simultaneously. The first method can check both suffix and longest common prefix~(LCP) arrays for strings drawn from constant alphabets, while the second method is applicable to checking SA for strings of any alphabets. By combining our methods with the Karp-Rabin fingerprinting function, we design two probablistic algorithms that perform verification with a neligible error probability. These algorithms are rather lightweight in terms of time and space. Particularly, the algorithm designed by the second method only takes linear time and constant space when running in external memory. For a further study, we integrate the checking algorithm designed by the second method into the existing suffix sorter DSA-IS to evaluate the checking overhead, where the time, space and I/O consumptions for verification are considerably smaller than that for construction in the experiments. This convinces us that the proposed checking methods, together with the state-of-the-art IS suffix sorters, could constitute efficient solutions for both construction and verification.

@@

The software packages for DC3 and eSAIS provide users a checker to ensure the correctness of their outputs. The following statements describe the idea behind this checker.

\begin{lemma} \label{lemma:1}
	For $i, j \in [0, n)$ and $i \ne j$, ${\sf suf}(i) < {\sf suf}(j) \Leftrightarrow (x[i], {\sf suf}(i + 1)) < (x[j], {\sf suf}(j + 1))$.
\end{lemma}

\begin{theorem} \label{theorem:1}
	$sa[0, n)$ is the SA for $x[0, n)$ if and only if the following conditions are satisfied:\\
	(1) $sa$ is a permutation of $[0, n)$. \\	
	(2) ${\sf r_i < r_j} \Leftrightarrow (x[i], r_{i + 1}) < (x[i], r_{j + 1})$ for $ i, j \in [0, n)$ and $i\ne j$, where $r_i$ and $r_j$ represent the ranks of ${\sf suf}(i)$ and ${\sf suf}(j)$ among all the suffixes, respectively. \\
\end{theorem}

\begin{IEEEproof}
	Theorem~\ref{theorem:1} is derived from Lemma~\ref{lemma:1}. The proof is available in~\cite{Dementiev2008a}.
\end{IEEEproof}

This checker performs two passes of integer sorts and each sort involves $n$ fixed-size tuples. When implemented in external memory, it consumes large disk space and I/O volume, resulting in a high checking overhead. For example, we can observe from Section~\ref{sec:experiments} that, the peak disk use and the I/O volume for checking an SA encoded by 40-bit integers are around $26n$ and $53n$, respectively. In this section,  we describe two SA checking methods based on the IS principle, both of them can be seamlessly integrated into any IS suffix sorting algorithm to build and check an SA at the same time.

\subsection{Method A} \label{sec:checkers:method_a}

Given that each S*-type substring is unique in $x$, SA-IS first computes $sa^*$ by sorting the starting positions of all the S*-type suffixes with their ranks indicated by $sa_1$ and then induces $sa$ from $sa^*$ by S\ref{induction_phase:1}''-S\ref{induction_phase:3}'' during the inducting phase.

\begin{enumerate}[S1'']

	\item Clear $sa$. Scan $sa^*$ leftward with $i$ decreasing from $n_1 - 1$ to $0$. For each scanned item $sa^*[i]$, insert it into the rightmost empty position of ${\sf sa\_bkt_S}(x[sa^*[i]])$. \label{induction_phase:1}

	\item Scan $sa$ rightward with $i$ increasing from $0$ to $n - 1$. For each scanned non-empty item $sa[i]$, insert $sa[i] - 1$ into the leftmost empty position of ${\sf sa\_bkt_L}(x[sa[i] - 1])$ if $t[sa[i] - 1] = 0$.\label{induction_phase:2}

	\item Clear ${\sf sa\_bkt_S}(c)$ for $c \in \Sigma$. Scan $sa$ leftward with $i$ decreasing from $n - 1$ to $0$. For each scanned non-empty item $sa[i]$, insert $sa[i] - 1$ into the rightmost empty position of ${\sf sa\_bkt_S}(x[sa[i] - 1])$ if $t[sa[i] - 1] = 1$.\label{induction_phase:3}

@@

We illustrate a running example of the above process in Fig.~\ref{fig:example1} and show more details for a better understanding. In line 6, we find the ending position of each bucket and insert the sorted S*-type suffixes into the right part of the corresponding buckets. Afterward, we first find the heading position of each bucket (marked by the symbol $\wedge$) and scan $sa$ rightward for inducing the order of L-type suffixes as following. In lines 8-9, we check $\sf suf(14)$ (marked by the symbol $@$) and find its predecessor ${\sf suf(13)}$ is L-type, thus we put ${\sf suf(13)}$ into the current leftmost empty position in ${\sf sa\_bkt_L}(i)$. To step through $sa$ in this way, we get all the L-type suffixes sorted in line 17. After that, we find the ending position of each bucket and scan $sa$ leftward for inducing the order of S-type suffixes in lines 18-28. When scanning $sa[14]$, we see $x[3]$ is S-type and thus put ${\sf suf(3)}$ into the current rightmost empty position in ${\sf sa\_bkt_S}(i)$. Following the same, we get all the S-type suffixes sorted in line 28.

@@

As stated below, S\ref{induction_phase:1}''-S\ref{induction_phase:3}'' also constitute the sufficient and necessary conditions for a correct $sa$.

\begin{lemma} \label{lemma:3}
	For any IS suffix sorting algorithm, its output $sa[0, n)$ is the SA for the input string $x[0, n)$ if and only if the following conditions are satisfied: \\ 
	(1) $sa^*$ is correct. \\
	(2) $sa[i]$ is equal to the value calculated by S\ref{induction_phase:1}''-S\ref{induction_phase:3}''. \\

@@

Suppose $sa^*$ is correct\footnote{The correctness of $sa^*$ can be checked by a sparse SA checker, such as~\cite{wu2017}. It should be noticed that the checking processes for $sa^*$ and $sa$ can be executed in parallel.}, this lemma suggests a method to check $sa$ when it is induced from $sa^*$ during the induction phase. Specifically, in S\ref{induction_phase:2}'' and S\ref{induction_phase:3}'', each item induced into $sa$ will be scanned later to induce the order of its predecessor. If the induced and scanned values for each item are equal, then $sa$ is correct according to Lemma~\ref{lemma:3}. The problem to be solved here is that when an item is induced into a bucket, its value in $sa$ will not be scanned at once. One solution for this is to record the induced and scanned values separately during the building process and check their equality afterward. Our solution is to integrate the building and checking processes into a whole, by ensuring the sequence of items induced into a bucket is identical to that scanned later. For the purpose, we can use a fingerprinting function to increasingly compute the fingerprints of both sequences and check their equality in constant time at the end of induction phase. If the two fingerprints for each bucket are equal, then the second condition of Lemma~\ref{lemma:3} will be seen with a high probability. As a result, $sa$ can be built and probabilistically checked at the same time. Following the previous discussion, we design a probabilistic algorithm based on  Corollary~\ref{corollary:1}.

\begin{corollary} \label{corollary:1}
	$sa[0, n)$ is the SA for the input string $x[0, n)$ with a high probability if the following conditions are satisfied for all $c \in \Sigma$: \\
	(1) $sa^*$ is correct.\\
	(2) The fingerprints of ${\sf sa_{I1}}(c)$ and ${\sf sa_{S1}}(c)$ are equal. \\
	(3) The fingerprints of ${\sf sa_{I2}}(c)$ and ${\sf sa_{S2}}(c)$ are equal. \\

\end{corollary}

Notice that ${\sf sa_{I1}}(c)$ and ${\sf sa_{I2}}(c)$ are two sequences respectively induced into ${\sf sa\_bkt_L}$ and ${\sf sa\_bkt_S}$, while ${\sf sa_{S1}}(c)$ and ${\sf sa_{S2}}(c)$ are two sequences respectively scanned from ${\sf sa\_bkt_L}$ and ${\sf sa\_bkt_S}$.

We point out that Method A can be also applied to checking $lcp$ when it is being built from $lcp^*$ following the IS principle~\cite{Fischer11}. In general, this is done by comparing the sequences of induced and scanned items for all the buckets in $lcp$.

\subsection{Method B}\label{sec:checkers:method_b}

{\color{red} 1. $\overline{sa^*}$ is a subset of the final $sa$, what is $sa^*[i]$? These two symbols have not been defined clearly and are confusing.

2. Notice: The whole algorithm for induced sorting suffixes consists of many processes, the inducing process is only one of them. The checking takes place only at the top level by reusing the inducing process there. For the checking algorithm to work, it must be correctly implemented, this implies that the inducing process at the top level is also implicitly assumed to be correctly implemented. This needs to be specified/assumed for clarity.
}

{ \color{blue} The definition of $sa^*$ was introduced in Section~\ref{sec:preliminaries}. An example was given on the left for better understanding.}

Our second Method is based on Theorem~\ref{theorem:2}, where $\overline{sa^*}$ is a subset of the final $sa$ for recording the starting positions of all the S*-type suffixes in $x$. {\color{blue}  We prove this theorem using Lemma~\ref{lemma:1} and Theorem~\ref{theorem:1}. 
\begin{theorem} \label{theorem:2}
	For any IS suffix sorting algorithm, its output $sa[0, n)$ is the SA for the input string $x[0, n)$ if and only if the following conditions are satisfied for all $i \in [0, n_1)$: \\
	(1) $sa^*[i]$ is a permutation of all the S*-type suffixes and  $sa^*[0]=n-1$. \\
	(2) $sa^*[i] = \overline{sa^*}[i]$. \\
\end{theorem}

\begin{IEEEproof}
	We only prove the sufficiency as the necessity is clear.

	Case 1: Suppose all the S*-type suffixes are correctly sorted, $sa$ is correct by Lemma \ref{lemma:3}.

	Case 2: {\color{red} Notice that the inducing process at the top level is correctly implemented by Assumption xxx.} 
	Suppose any two S*-type suffixes are not correctly sorted, i.e. ${\sf suf}(sa^*[i_0])>{\sf suf}(sa^*[j_0])$ and $i_0<j_0$. By condition (2), we have ${\sf suf}(\overline{sa^*}[i_0])>{\sf suf}(\overline{sa^*}[j_0])$. Because the order of ${\sf suf}(\overline{sa^*}[i_0])$ and ${\sf suf}(\overline{sa^*}[j_0])$ is induced from ${\sf suf}(sa^*[i_1])$ and ${\sf suf}(sa^*[j_1])$, where ${\sf sub}(\overline{sa^*}[i_0], sa^*[i_1])$ and ${\sf sub}(\overline{sa^*}[j_0], sa^*[j_1])$ are two LMS-substrings, there must be ${\sf suf}(sa^*[i_1])>{\sf suf}(sa^*[j_1])$ and $i_1<j_1$. Repeating this reasoning process, because condition (1), we must see ${\sf suf}(sa^*[i_k])>{\sf suf}(sa^*[j_k])$ and ${\sf suf}(sa^*[i_{k+1}])<{\sf suf}(sa^*[j_{k+1}])$, where $i_k<j_k$ and $i_{k+1}<j_{k+1}$. However, given ${\sf suf}(sa^*[i_{k+1}])<{\sf suf}(sa^*[j_{k+1}])$, the inducing process will produce ${\sf suf}(\overline{sa^*}[i_k])<{\sf suf}(\overline{sa^*}[j_k])$ , which implies ${\sf suf}(sa^*[i_k])<{\sf suf}(sa^*[j_k])$ because condition (2), leading to a contradiction.
\end{IEEEproof}

The first condition of Theorem~\ref{theorem:2} is naturally satisfied when computing $sa^*$ from $sa_1$. A naive solution for checking the second condition is to keep a copy of $\overline{sa^*}$ when computing $sa$ from $sa^*$ and compare it with $sa^*$ afterward. This takes linear time, space and I/O overhead. We prefer to check equality of them by comparing their fingerprints, which can be calculated during the scan of these arrays in S\ref{induction_phase:1}'' and S\ref{induction_phase:3}''.

\begin{corollary} \label{corollary:2}
	$sa[0, n)$ is the SA for the input string $x[0, n)$ with a high probability if the following conditions are satisfied for all $i \in [0, n_1)$: \\
	(1) $sa^*[i]$ is a permutation of all the S*-type suffixes and  $sa^*[0]=n-1$. \\
	(2) The fingerprints of $sa^*$ and $\overline{sa^*}$ are equal. \\
\end{corollary}

As can be seen in Fig.~\ref{fig:example1}, both $sa^*$ and $\overline{sa^*}$ contain items $\{14, 8, 5, 2\}$ arranged in the same order. By using the fingerprinting function introduced in the next subsection, the fingerprints can be calculated in linear time using constant space. This implies another probabilistic algorithm designed by Method B.

@@

@@

	\includegraphics[width = 0.9\columnwidth]{example2.pdf}

@@

Methods A and B check equality of two integer arrays by computing and comparing their fingerprints. In our programs, we select the Karp-Rabin fingerprinting function to compute the fingerprints in need. As depicted in Fig.~\ref{fig:example2}, the fingerprint of $A$ is calculated according to Formulas~\ref{formula:1}-\ref{formula:2}, where $L$ is a prime and $\delta$ is an integer randomly chosen from $[1, L)$. It should be noticed that two equal arrays must have an identical fingerprint, but the inverse is not always true. Fortunately, the error probability can be reduced to a negligible level by setting $L$ to a large value.

>>>>

Suffix array~(SA) can be built by the induced sorting~(IS) method on both internal and external memory models. We propose two methods that enable any IS suffix sorter to build and check an SA simultaneously. The first method can check both suffix and longest common prefix~(LCP) arrays, while the second method is applicable to checking SA only. By combining our methods with the Karp-Rabin fingerprinting function, we design two probablistic algorithms that perform verification with a neligible error probability. These algorithms are rather lightweight in terms of time and space. Particularly, the algorithm designed by the second method only takes linear time and constant space when running in external memory. For a further study, we integrate the checking algorithm designed by the second method into the existing suffix sorting algorithm DSA-IS to evaluate the checking overhead, where the time, space and I/O consumptions for verification are considerably smaller than that for construction in the experiments. This convinces us that the proposed checking methods, together with the state-of-the-art IS suffix sorters, could constitute efficient solutions for both construction and verification.

@@

\color{red}{The software packages for DC3 and eSAIS provide users a checker to ensure the correctness of their outputs. Below we describe the main idea behind this checker and refer interested readers to ~\cite{Dementiev2008a} for more details.}

\begin{lemma} \label{lemma:1}
	$sa[0, n)$ is the SA for $x[0, n)$ if and only if the following conditions are satisfied:\\
	(1) $sa$ is a permutation of $[0, n)$. \\	
	(2) ${\sf r_i < r_j} \Leftrightarrow (x[i], r_{i + 1}) < (x[i], r_{j + 1})$ for $ i, j \in [0, n)$ and $i\ne j$, where $r_i$ and $r_j$ represent the ranks of ${\sf suf}(i)$ and ${\sf suf}(j)$ among all the suffixes, respectively. \\
\end{lemma}

\begin{IEEEproof}The first condition indicates that $sa$ is a permutation of all the suffixes in $x$. The second condition indicates that the order of suffixes in $sa$ corresponds to that of their heading characters and successors. Because a comparison between any two suffixes can be done by comparing their heading characters and successors, the above conditions are sufficient and necessary for SA verification.

\end{IEEEproof}

The disk-based implementation of this checker conducts two passes of integer sorts and each sort arranges the order of $n$ fixed-size tuples using external memory. As can be observed from Section~\ref{sec:experiments}, the peak disk use and the I/O volume for an SA encoded by 40-bit integers are around $26n$ and $53n$, respectively.

\subsection{Proposals} \label{sec:checkers:proposals}

Recall that, an IS suffix sorting algorithm recursively perform the reduction phase to compute the reduced string $s_1$ until $s_1$ contains no duplicate characters. Afterward, it produces $sa_1$ from $s_1$ and recursively perform the induction phase to compute $sa$ until reaching the top recursion level, where the induction phase executes mainly consists of the following steps:

\begin{enumerate}[S1'']
	\item sort the starting positions of all the S*-type suffixes with their ranks indicated by $sa_1$ to produce $sa^*$ \label{induction_phase:1}.

	\item Clear $sa$. Scan $sa^*$ leftward with $i$ decreasing from $n_1 - 1$ to $0$. For each scanned item $sa^*[i]$, insert it into the rightmost empty position of ${\sf sa\_bkt_S}(x[sa^*[i]])$. \label{induction_phase:2}

	\item Scan $sa$ rightward with $i$ increasing from $0$ to $n - 1$. For each scanned non-empty item $sa[i]$, insert $sa[i] - 1$ into the leftmost empty position of ${\sf sa\_bkt_L}(x[sa[i] - 1])$ if $t[sa[i] - 1] = 0$.\label{induction_phase:3}

	\item Clear ${\sf sa\_bkt_S}(c)$ for $c \in \Sigma$. Scan $sa$ leftward with $i$ decreasing from $n - 1$ to $0$. For each scanned non-empty item $sa[i]$, insert $sa[i] - 1$ into the rightmost empty position of ${\sf sa\_bkt_S}(x[sa[i] - 1])$ if $t[sa[i] - 1] = 1$.\label{induction_phase:4}

@@

A running example of S\ref{induction_phase:2}-Sp\ref{induction_phase:4} is shown in Fig.~\ref{fig:example1}. Given $sa^*$ is already known, line 6 inserts each S*-type suffix into the right part of the corresponding bucket and arranges them according to their sorted order indicated by $sa^*$. For example, $\sf suf(8)$, $\sf suf(5)$, $\sf suf(2)$ are placed into the right part of $\sf sa\_bkt(i)$ and arranged in their sorted order indicated by $sa^*$. Then, we find the leftmost position of each bucket (marked by the symbol $\wedge$) and scan $sa$ rightward for inducing the order of L-type suffixes. To do this, we first check $sa[0] = 14$ (marked by the symbol @) and induce the predecessor of $\sf suf(14)$ in lines 8-9. Because $x[13] = i$ is L-type, we put $\sf suf(13)$ into the current leftmost empty position in $\sf sa\_bkt_L(i)$. To step through $sa$ in this way, we get all the L-type suffixes sorted in line 17. Afterward, we find the rightmost position of each bucket and scan $sa$ leftward for inducing the order of S-type suffixes. When scanning $sa[14] = 4$ in lines 19-20, we see $x[3] = i$ is S-type and thus put $\sf suf(3)$ into the current rightmost empty position in $\sf sa\_bkt_S(i)$. Following the same idea, we get all the S-type suffixes sorted in line 28. Based on the discussion, we show in Lemma~\ref{lemma:2} a set of sufficient and necessary conditions for SA verification.

@@

\begin{lemma} \label{lemma:2}
    $sa[0, n)$ is the SA for $x[0, n)$ if and only if the following conditions are satisfied: \\
    (1) $sa^*$ is correctly computed. \\
    (2) S\ref{induction_phase:1}-S\ref{induction_phase:4} are correctly implemented. \\
\end{lemma}

\begin{IEEEproof} Given two suffixes placed at $sa[i]$ and $sa[j]$ and their successors placed at $sa[p]$ and $sa[q]$, we prove the statement as follows.

Because S\ref{induction_phase:1}-S\ref{induction_phase:4} are correctly implemented, $i < j \iff (x[sa[i]], p) < (x[sa[j]], q)$. This satisfies the second condition of Lemma~\ref{lemma:1}.

Further, suppose $sa[i] = sa[j]$ and $i \ne j$, then $sa[p] = sa[q]$ and $p \ne q$. Repeatedly replacing $(i, j)$ with $(p, q)$ until $\sf suf(sa[i])$ is S*-type, then $sa^*$ contain duplicate elements. Because $sa^*$ is correctly computed, each element in $sa^*$ must be unique. This leads to a contradiction.
\end{IEEEproof}

The above lemma indicates that, if the first condition of Lemma~\ref{lemma:2} is true, we can check the correctness of $sa^*$ instead of $sa$ at the top recursion level to perform SA verification. It should be noticed that the code snippet for the induction phase at the top recursion level only consists of tens of C++ code lines. This is considerable smaller than the whole program for a disk-based suffix sorting algorithm. Following the idea, we assume S\ref{induction_phase:1}-S\ref{induction_phase:4} are correctly implemented and propose two methods for checking computation errors caused by implementation bugs and I/O errors caused by hardware malfunctions. The algorithms designed by these methods can be seamlessly integrated into any IS suffix sorting algorithms for building and checking an SA at the same time.

\subsubsection{Method A} \label{sec:proposals:method_a}

Our first method is directly extended from Lemma~\ref{lemma:2}.

\begin{lemma} \label{lemma:3}
	Assume the induction phase at the top recursion level is correctly implemented, the output $sa[0, n)$ of an IS suffix sorting algorithm is the SA for the input string $x[0, n)$ if and only if the following conditions are satisfied: \\
	(1) $sa^*$ is correct. \\
	(2) $sa[i]$ is equal to the value calculated by S\ref{induction_phase:1}''-S\ref{induction_phase:4}''. \\

@@

The first condition of Lemma~\ref{lemma:3} can be checked by a sparse SA checker like~\cite{wu2017}. Notice that each suffix induced into $sa$ will be latter scanned for inducing the order of its predecessor during S\ref{induction_phase:2}'' and S\ref{induction_phase:4}''. The second condition intends for detecting I/O errors and can be checked by determining whether the induced and scanned values for each suffix are equal. The problem to be solved here is that when a suffix is induced into a bucket, its corresponding value in $sa$ may not be scanned at once. Our solution is to integrate the checking process into the building process, by ensuring the sequence of values placed at a bucket is identical to that scanned later. For the purpose, we use a fingerprinting function to increasingly compute the fingerprints of both sequences and check their equality in constant time at the end of the induction phase. If the two fingerprints for each bucket are equal, then the second condition of Lemma~\ref{lemma:3} will be seen with a high probability. As a result, $sa$ can be built and probabilistically checked at the same time. It should be noticed that this method can be also applied to checking the longest common prefix array when it is being built from $lcp^*$ following the IS principle~\cite{Fischer11}. As a result, we design Algorithm~\ref{alg:2} to probabilistically check $sa$ and $lcp$ during the induction phase at the top recursion level.

\SetKwProg{Fn}{Function}{}{}

\begin{algorithm*}

	\caption{The Algorithm Based on Lemma~\ref{lemma:3}.}

	\label{alg:2}

	%\SetAlgoNoLine
	\Fn{{\sf CheckByMethodA}}{

        Check the correctness of $sa^*$ and $lcp^*$ by the method proposed in~\cite{wu2017}. \\

        Compute the fingerprints of ${\sf sa_{I1}}(c)$, ${\sf sa_{S1}}(c)$, ${\sf lcp_{I1}}(c)$ and ${\sf lcp_{S1}}(c)$ when inducing the order and the LCP-values of L-type suffixes. \\

        Check if ${\sf sa_{I1}}(c) = {\sf sa_{S1}}(c)$ and ${\sf lcp_{I1}}(c) = {\sf lcp_{S1}}(c)$. \\

        Compute the fingerprints of ${\sf sa_{I2}}(c)$, ${\sf sa_{S2}}(c)$, ${\sf lcp_{I2}}(c)$ and ${\sf lcp_{S2}}(c)$ when inducing the order and the LCP-values of S-type suffixes. \\

        Check if ${\sf sa_{I2}}(c) = {\sf sa_{S2}}(c)$, ${\sf lcp_{I2}}(c) = {\sf lcp_{S2}}(c)$. \\
	}
	\end{algorithm*}

\subsubsection{Method B}\label{sec:proposals:method_b}

Method B uses a different way to check the correctness of $sa^*$ in Lemma~\ref{lemma:4}. Before our presentation, we first introduce a notation called $\overline{sa^*}$. The same as $sa^*$, $\overline{sa^*}$ also represents the suffix array for all the S*-type suffixes. Both of them are computed during the induction phase at the top recursion level of an IS suffix sorting algorithm. To be specific, $sa^*$ is computed from $sa_1$ and $\overline{sa^*}$ is retrieved from the constructed $sa$.

{\color{red}
\begin{lemma} \label{lemma:4}
    Assume the induction phase at the top recursion level is correctly implemented, the output $sa[0, n)$ of an IS suffix sorting algorithm is the SA for the input string $x[0, n)$ if and only if the following conditions are satisfied for all $i \in [0, n_1)$:

	(1) $sa^*[i]$ is a permutation of all the S*-type suffixes. \\
	(2) $sa^*[i] = \overline{sa^*}[i]$. \\
	(3) $sa[i]$ is equal to the value calculated by S\ref{induction_phase:1}''-S\ref{induction_phase:4}''. \\

\begin{IEEEproof}

    We only prove the sufficiency as the necessity is clear.

    Case 1: Suppose all the S*-type suffixes are correctly sorted in $sa^*$, then $sa$ is right according to Lemma~\ref{lemma:3}.

    Case 2: Suppose any two S*-type suffixes are not corrected sorted, i.e., $\sf suf(sa^*[i_0]) > \sf suf(sa^*[j_0])$ and $i_0 < j_0$. By condition (2), we have $\sf suf(\overline{sa^*}[i_0] > \sf suf(\overline{sa^*}[j_0])$. Given that the order of $\sf suf(\overline{sa^*}[i_0]$ and $\sf suf(\overline{sa^*}[j_0])$ are induced from ${\sf suf}(sa^*[i_1])$ and ${\sf suf}(sa^*[j_1])$, then ${\sf sub}(\overline{sa^*}[i_0], sa^*[i_1])$ and ${\sf sub}(\overline{sa^*}[j_0], sa^*[j_1])$ are two LMS-substrings and there must be ${\sf suf}(sa^*[i_1])>{\sf suf}(sa^*[j_1])$ and $i_1 < j_1$. Repeating this reasoning process, because condition (1), we must see ${\sf suf}(sa^*[i_k])>{\sf suf}(sa^*[j_k])$ and ${\sf suf}(sa^*[i_{k+1}])<{\sf suf}(sa^*[j_{k+1}])$, where $i_k < j_k$ and $i_{k+1} < j_{k+1}$. However, given ${\sf suf}(sa^*[i_{k+1}])<{\sf suf}(sa^*[j_{k+1}])$, the inducing process will produce ${\sf suf}(\overline{sa^*}[i_k])<{\sf suf}(\overline{sa^*}[j_k])$ , which implies ${\sf suf}(sa^*[i_k])<{\sf suf}(sa^*[j_k])$ because condition (2), leading to a contradiction.

\end{IEEEproof}
}

The first condition of Lemma~\ref{lemma:4} is naturally satisfied when computing $sa^*$ from $sa_1$. We reuse the fingerprinting function to probabilistically check the third condition. This method is also applied to checking the second condition by computing and comparing the fingerprints of $sa^*$ and $\overline{sa^*}$. Specifically, the fingerprint of $sa^*$ is calculated when it is being scanned in S$\ref{induction_phase:1}''$. Because $\overline{sa^*}$ is a subset of $sa$, the fingerprint of $\overline{sa^*}$ is calculated when the S-type suffixes are being induced in S$\ref{induction_phase:4}''$.

\SetKwProg{Fn}{Function}{}{}

\begin{algorithm*}

	\caption{The Algorithm Based on Lemma~\ref{lemma:4}.}

	\label{alg:3}

	%\SetAlgoNoLine
	\Fn{{\sf CheckByMethodB}}{

        Compute the fingerprints of $sa^*$, ${\sf sa_{I1}}(c)$ and ${\sf sa_{S1}}(c)$ when inducing the order and the LCP-values of L-type suffixes. \\

        Check if ${\sf sa_{I1}}(c) = {\sf sa_{S1}}(c)$ and ${\sf lcp_{I1}}(c) = {\sf lcp_{S1}}(c)$ by comparing their fingerprints. \\

        Compute the fingerprints of $\overline{sa^*}$, ${\sf sa_{I2}}(c)$ and ${\sf sa_{S2}}(c)$ when inducing the order and the LCP-values of S-type suffixes. \\

        Check if ${\sf sa_{I2}}(c) = {\sf sa_{S2}}(c)$p by comparing their fingerprints. \\

        Check if $sa^* = \overline{sa^*}$ by comparing their fingerprints. \\
	}
\end{algorithm*}

@@

The proposed two methods check equality of two arrays by comparing their fingerprints, which can be calculated using a hash function. Notice that two equal arrays must share an identical hash value, but the inverse is not always true. To lower the error probability of a false match, we prefer using the Karp-Rabin fingerprinting function to compute the required values following Formulas~\ref{formula:1}-\ref{formula:2}, where $L$ is a prime and $delta$ is an integer randomly chosen from $[1, L)$. By setting $L$ to a large value, the error probability can be reduced to a negligible level. In Fig.~\ref{fig:example2}, we depict an example for computing the integer array $A$, where $A$ is identical to $sa^*$ and $\overline{sa^*}$ in Fig.~\ref{fig:example1}. Given $L = 23$ and $\delta = 5$, we first compute the fingerprint of $A[0, 0]$ in line 4. Because $\sf fp(A[0, -1]) = 0$, $\sf fp(A[0, 0]) = (0 \cdot 5 + 14) \, mod \, 23 = 14$. By iteratively computing the fingerprints of the prefixes of $A$ according to Formula~\ref{formula:2}, we finally obtain the fingerprints of the whole array in line 7.

@@

	\includegraphics[width = 1\columnwidth]{example2.pdf}

@@

\subsection{Complexity Analysis}

It takes linear time and constant RAM space for computing the required fingerprints in Algorithm~\ref{alg:2} and Algorithm~\ref{alg:3}. Further, the bottleneck of Algorithm~\ref{alg:2} lies in checking $sa^*$ following the idea presented in~\cite{wu2017}. Specifically, given two suffixes, say $\sf suf(i)$ and $\sf suf(j)$, and their LCP-value, say $d$, we compute and compare the fingerprints of $\sf sub(i, i + d - 1)$ and $\sf sub(j, j + d - 1)$. If the fingerprints are equal and the order of $x[i + d]$ and $x[j + d]$ correspond to that of $\sf suf(i)$ and $\sf suf(j)$, it is believed that the suffixes are correctly sorted and their LCP-value is right. When implemented in external memory, the checking processes for any number of suffix pairs can be done in batch within a sorting complexity. Because at most one out of every two suffixes is S*-type (typically one-third in real-world data sets) and the checking process is only executed at the top recursion level, the checking consumption taken by Algorithm~\ref{alg:2} is much less than the building consumption taken by any disk-based IS suffix sorting algorithm.
